---
title: "Solutions R for Dataseciece"
author: "Nagesh Bheesetty"
output: rmarkdown::github_document
---

CS 625, Fall 2020








### Goal:

To find solutions to all the exercise questions in the book "R for Data Science".





## Report:


**Importing the libraries "tidyverse", and "lubridate"**.  


```{r, message = FALSE, echo=FALSE}

library(tidyverse)
library(dplyr)
library(lubridate)


```

The library "**tidyverse**" allows statisticians and data scientists to create work-flow that facilitate communication and result in reproducible work product. It consists a collection of R packages that share an underlying design philosophy, grammar, and data structures. this "tidyverse" makes data processing faster and easier.
The most important packages are:

  * ggplot
  * dplyr
  * tidyr
  * readr
The library "**lubridate**" allowed me to edit the date format in the datasets.  


## Source:

[R for Data Science](https://r4ds.had.co.nz/data-visualisation.html)

### 3 Data visualisation:



#### 3.2.1 The mpg data frame

```{r}
mpg

```

Among the variables in `mpg` are:

`displ`, a car’s engine size, in litres.

`hwy`, a car’s fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance.

To learn more about `mpg`, open its help page by running `?mpg`.


### 3.2.2 Creating a ggplot

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### 3.2.3 A graphing template

```{}
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

```

### 3.2.4 Exercises

1. Run `ggplot(data = mpg)`. What do you see?

2. How many rows are in `mpg`? How many columns?

3. What does the `drv` variable describe? Read the help for `?mpg` to find out.

4. Make a scatterplot of `hwy` vs `cyl`.

5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?


#### Solutions 3.2.4

**1. Run ggplot(data = mpg). What do you see?**

```{r}

ggplot(data = mpg)
```
Just by running `ggplot(data = mpg)` we could not see any output but only a plot. However, the `ggplot(data = mpg)` must be added with the `geom` functions.


**2. How many rows are in `mpg`? How many columns?**

```{r}
mpg
```

By running `mpg` we can find the number of rows and columns.

It is observed that there are 234 rows and 11 columns.

To find the the number of rows and columns with out displaying the data, we can use `nrow()` and `ncol()`.


```{r}
# Number of rows in `mpg`
nrow(mpg)
# Number of columns in `mpg`
ncol(mpg)
```
The `glimpse()` function also displays the number of rows and columns in a data frame

```{r}
glimpse(mpg)
```


**3. What does the `drv` variable describe? Read the help for `?mpg` to find out.**

```{r}
?mpg
```

|Value|Description|
|---|---|
|drv|the type of driven train|
|"f"|front-wheel drive|
|"r"|rear-wheel drive|
|"4"| 4wd|


**4. Make a scatterplot of `hwy` vs `cyl`**.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```



**5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?**

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))

```

Scatterplot can basically establish relation between two quantitative variables to find the trend and relation between the variables. In this case `class` and `drv` are both categorical variables. So, the plot is not useful. 

```{r}
count(mpg, drv, class)
```

```{r}
ggplot(mpg, aes(x = class, y = drv, color = class)) + 
    geom_count()
```


### 3.3 Aesthetic mappings

Color is a good idea to distinguish the categorical types within the values.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
#> Warning: Using size for a discrete variable is not advised.
```








**Line Graph for COVID-19:**

```{ echo=FALSE}
COVID <- read.csv("./HW7_datasets/COVID-19.csv")
#head(COVID)
COVID$date<-mdy(COVID$date)
COVID19 <- select(COVID, date, cases)
COVID19 <- rename(COVID19, Date = date)
# #head(COVID19)
COVID19<- mutate(COVID19, Percentage = cases/sum(cases))
head(COVID19)
ggplot(data = COVID19, mapping = aes(x = Date, y = Percentage)) +
  geom_line(size=1.5, color="red")+
  geom_point(size=0.5, color="blue")+labs(x="Jan-2020 to Nov-2020", y ="Number of cases" , title = "COVID-19 cases from Jan-2020 through Nov-2020")
```


**Line Graph for Gold Stock:**

```{, echo=FALSE}
Gold<-read.csv("./HW7_datasets/gold_yahoo.csv")
#head(Gold)

Gold$Date<-as.Date(Gold$Date)

Gold <- select(Gold, Date, Open)
Gold <- rename(Gold, Price = Open)
#head(Gold)
Gold<- mutate(Gold, Percentage = Price*100/sum(Price))
Gold = filter(Gold, Date > as.Date("2019-11-01","%Y-%m-%d"))
#head(Gold)

ggplot(data = Gold, mapping = aes(x = Date, y = Price)) +

  geom_point(size=1.5, color="blue") + geom_line(size = 0.5, color = "green")+theme(axis.text.x = element_text(angle=45, hjust = 1))+labs(x="Jan-2019 to Nov-2020", y ="Stock price in dollars" , title = "Gold stock price from Jan-2019 through Nov-2020")

```

**Line Graph for Gas Prices:**



```{, echo= FALSE}
Gas<-read.csv("./HW7_datasets/gas_monthly_csv.csv")
#head(Gas)
Gas$Date = as.Date(paste(Gas$Month,"-01",sep=""))
Gas = filter(Gas, Date > as.Date("2019-11-01","%Y-%m-%d"))
#head(Gas)
Gas <- select(Gas, Date, Price)
#Gold <- rename(Gold, Price = Open)
#head(Gas)
Gas<- mutate(Gas, Percentage = Price*100/sum(Price))
#head(Gas)

ggplot(data = Gas, mapping = aes(x = Date, y = Price)) +

  geom_line(size=1, color="purple")+
  geom_point(size=1.5, color="blue")+theme(axis.text.x = element_text(angle=45, hjust = 1))+labs(x="Jan-2019 to Nov-2020", y ="Gas price in dollars" , title = "Gas price from Jan-2019 through Nov-2020")

```

**Line Graph for SP500:**

```{, echo= FALSE}
SP500<-read.csv("./HW7_datasets/S&P500_yahoo.csv")
#head(SP500)
SP500$Date = as.Date(SP500$Date)
#head(SP500)
SP500 <- select(SP500, Date, Open)
SP500 <- rename(SP500, Price = Open)
#head(SP500)
SP500<- mutate(SP500, Percentage = Price*100/sum(Price))
#head(SP500)
SP500 = filter(SP500, Date > as.Date("2019-11-01","%Y-%m-%d"))
#head(SP500)
ggplot(data = SP500, mapping = aes(x = Date, y = Price)) +

  geom_line(size=1, color="orange")+
  geom_point(size=1.5, color="blue")+

  theme(axis.text.x = element_text(angle=45, hjust = 1))+labs(x="Jan-2019 to Nov-2020", y ="Stock price in dollars" , title = "S&P500 stock price from Jan-2019 through Nov-2020")

```






















**Procedures and Design Decisions**

|Idiom:| Smallmultiple line graphs|
|---|---|
|X-Axis:| Jan-2020 to Nov-2020 (first letters of each month)| 
|Y-Axis(left):| COVID (Number of cases)|
|Y-Axis(Right):| Comparison Attributes (Unemployment, Gas, Stocks)|


### Explanation:









### References:


**Change the Date format:** [STATMETHODS](https://www.statmethods.net/input/dates.html)

**Add Colors to each line:** [Colors in R]()

**How to Export DataFrame to CSV in R:** [Data to Fish](https://datatofish.com/export-dataframe-to-csv-in-r/)

**library(tidyverse):** https://ggplot2.tidyverse.org/reference/index.html

**Split the date:** [Stack overflow](https://stackoverflow.com/questions/4310326/convert-character-to-date-in-r)

**Exploratory Data Analysis:(EDA)** [R for Data Science](https://r4ds.had.co.nz/)















