---
title: "Solutions R for Dataseciece"
author: "Nagesh Bheesetty"
output: rmarkdown::github_document
---

CS 625, Fall 2020








### Goal:

To find solutions to all the exercise questions in the book "**R for Data Science**".





## Report:


**Importing the libraries "tidyverse", and "lubridate"**.  


```{r, message = FALSE, echo=FALSE}

library(tidyverse)
library(dplyr)
library(lubridate)


```

The library "**tidyverse**" allows statisticians and data scientists to create work-flow that facilitate communication and result in reproducible work product. It consists a collection of R packages that share an underlying design philosophy, grammar, and data structures. this "tidyverse" makes data processing faster and easier.
The most important packages are:

  * ggplot
  * dplyr
  * tidyr
  * readr
The library "**lubridate**" allowed me to edit the date format in the datasets.  


## Source:

[R for Data Science](https://r4ds.had.co.nz/data-visualisation.html)

### Chapter-3:(Data visualisation)



#### 3.2.1 The `mpg` data frame

```{r}
mpg

```

Among the variables in `mpg` are:

`displ`, a car’s engine size, in litres.

`hwy`, a car’s fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance.

To learn more about `mpg`, open its help page by running `?mpg`.


### 3.2.2 Creating a ggplot

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### 3.2.3 A graphing template

```{}
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

```

### 3.2.4 Exercises

1. Run `ggplot(data = mpg)`. What do you see?

2. How many rows are in `mpg`? How many columns?

3. What does the `drv` variable describe? Read the help for `?mpg` to find out.

4. Make a scatterplot of `hwy` vs `cyl`.

5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?


#### Solutions 3.2.4

**1. Run `ggplot(data = mpg)`. What do you see?**

```{r}

ggplot(data = mpg)
```
Just by running `ggplot(data = mpg)` we could not see any output, except a plot.
Because, `ggplot()` is not instructed what to plot. It means we should define a plot between two `variables` (attributes in the columns). So, the `ggplot(data = mpg)` must be added with the `geom` functions.


**2. How many rows are in `mpg`? How many columns?**

There are multiple ways to find the number fo rows and columns in the data frame.

 * Just by giving the data (say `mpg`) we can find the number of rows and columns, but this will run the whole data frame as well. That means a lot of bites and space.

```{r}
mpg
```

It is observed that there are 234 rows and 11 columns.

 * The best way to find the the number of `rows` and `columns` with out displaying the data frame is to use commands `nrow()` and `ncol()`.


```{r}
# Number of rows in `mpg`
nrow(mpg)
# Number of columns in `mpg`
ncol(mpg)
```
This will display the number of rows and columns.

1 Rows: 234

2 Columns: 11


* The alternate way is  `glimpse()` function which can also display the number of rows and columns in a data frame

```{r}
glimpse(mpg)
```
This function `glimpse()` is good to find all the columns in a row and the attribute types can be easily identified.

In the above output it can be observed that out of the 11 columns 6 attributes are categorical and 5 of them are quantitative.

Assume a data frame with 20, 40, or more number of columns all the attribute types can be identified.


**3. What does the `drv` variable describe? Read the help for `?mpg` to find out.**

```{r}
?mpg
```

|Value|Description|
|---|---|
|drv|the type of driven train|
|"f"|front-wheel drive|
|"r"|rear-wheel drive|
|"4"| 4wd|

Most of the times we see a data attributes with short words or abbreviations and we may also see some short form of categorical information in the columns. 

By running `?data` say `?mpg` the information about the data can be identified.

I believe the predefined packages in _ggplot_ has the description about the data.


**4. Make a scatterplot of `hwy` vs `cyl`**.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```

A scatterplot is a point mapping between two quantitative variables.

It is very important to ask a question in how to establish a relation, in general we tend to establish relation between one quantitative attribute varies with respect to the other. 

For example, as x changes there could be a change in y. So, x is placed on x-axis and y is placed on y-axis.

In the above question the task is to establish a relation between `hwy` and `cyl`. So, let us understand the attributes:

hwy $-->$ highway miles per gallon

cyl $-->$ number of cylinders

Might the two given attributes may or may not be correlated, however we can make a question and clarify that as the number of cylinders changes in the cars, the miles per gallon could vary but by increasing the miles per gallon we can not chane the number of cylinders in the car.

So, on the x-axis we need `cyl` and on the y-axis we can place `hwy`.

Lets run the code by changing the attributes.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(y = hwy, x = cyl))
```

However, it is evident that there is no relation between `cyl` and `hwy`.


**5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?**

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))

```

Scatterplot can basically establish relation between two quantitative variables to find the trend and relation between the variables. In this case `class` and `drv` are both categorical variables. So, the plot is not useful. 

This can be more clearly observed from the concepts in other chapters.


```{r}
count(mpg, drv, class)
```

```{r}
ggplot(mpg, aes(x = class, y = drv, color = class)) + 
    geom_count()
```


### 3.3 Aesthetic mappings

Color is a good idea to distinguish the categorical types within the values.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
#> Warning: Using size for a discrete variable is not advised.
```

















### References:


**Change the Date format:** [STATMETHODS](https://www.statmethods.net/input/dates.html)

**Add Colors to each line:** [Colors in R]()

**How to Export DataFrame to CSV in R:** [Data to Fish](https://datatofish.com/export-dataframe-to-csv-in-r/)

**library(tidyverse):** https://ggplot2.tidyverse.org/reference/index.html

**Split the date:** [Stack overflow](https://stackoverflow.com/questions/4310326/convert-character-to-date-in-r)

**Exploratory Data Analysis:(EDA)** [R for Data Science](https://r4ds.had.co.nz/)















